---
title: "Foreasting Australia's beer production"
author: 'Kyle Schmidt'
date: '3/15/2020'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install the necessary packages.
```{r message=FALSE}
library(TSA)
library(forecast)
```

Import Beer Production in Australia (in Megaliters). Data can be found at this link: https://www.kaggle.com/shenba/time-series-datasets#monthly-beer-production-in-austr.csv
```{r}
beer <- read.csv('monthly-beer-production-in-austr.csv')
```

Create time-series object and initial time-series plot.
```{r}
# Create time-series object
ts.beer <- ts(beer$Monthly.beer.production, start=c(1956, 1), deltat=1/12)

# Plot
plot(ts.beer, main='Monthly Beer Production in Australia', xlab='Year', ylab='Beer (Megaliters)')
```

The variabilty of the data increases over time. A log transform will stabilize the data.
```{r}
# Log Tranform
log.beer <- log(ts.beer)

# Plot
plot(log.beer, main='Monthly Beer Production in Australia', xlab='Year', ylab='Beer (Log Megaliters)')
```

The log of the monthly beer production time-series plot shows an increasing trend. Estimate the trend.
```{r}
# Create a time veriable
time.beer <- time(log.beer)

# Fit a quadratric model
lm.beer <- lm(log.beer~time.beer + I(time.beer^2))

# Estimate the trend using the linear model
pred.beer <-predict(lm.beer, newdata=time.beer)

# Create time-series object from the estimated trend
trend.beer <- ts(pred.beer, start=c(1956, 1), deltat=1/12)

# Plot the trend
plot(log.beer, main='Monthly Beer Production in Australia', xlab='Year', ylab='Beer (Log Megaliters)')

# Add trend line to the plot
lines(trend.beer, col = "red")
```

Remove the estimated trend from the log time-series by plotting the residuals.
```{r}
# Change the residuals into a time-series object
trend.res <- ts(lm.beer$residuals, start=c(1956, 1), deltat=1/12)

# plot the residuals
plot(trend.res, main='Trend Adjusted Log Monthly Beer Production in Australia', xlab='Year', ylab='Beer (Log Megaliters)')
```

There is also seasonality in the time-series. Estimate the seasonality.
```{r}
# Extract the month from each point
month <- cycle(log.beer)

# Calculate the mean of each month
res.lm <- lm(trend.res~factor(month))

# Subtract each point by the corresponding monthly mean
season.beer <- ts(res.lm$fitted.values, start=c(1956, 1), deltat=1/12)

# plot the seasonality
plot(season.beer, main='Seasonal Beer Production in Australia', xlab ='Year', ylab='Beer (Log Megaliters)')
```

Remove the seasonality from the trend adjusted log monthly beer production.
```{r}
# remove the seasonality
rand.beer <- ts(residuals(res.lm), start=c(1956, 1), deltat = 1/12)

# plot the final series
plot(rand.beer, main='Trend and Seasonality Adjusted Monthly Beer Production in Australia', xlab='Year', ylab='Beer (Log Megaliters)')
```

Alternatively trend and seasoanlity can be removed by differencing.
```{r}
dif.beer <- diff(log.beer)

plot(dif.beer, main='Differenced Log Beer Production in Australia', xlab='Year', ylab='Difference')
```

The differenced time-series looks to be closer to stationary than the manual trend and seasonality adjusted time-series. Look at the autocorrelation and partial autcorrelation functions of the differenced time-series to fit a model.
```{r}
# ACF plot
acf(dif.beer, lag.max=48, main = 'ACF for Differenced Series')

# PACF plot
pacf(dif.beer, lag.max=48, main = 'PACF for Differenced Series')
```

The initial ACF and PACF plot indicates seasonality each 12 months. Use a seasonal difference at lag 12.
```{r}
seasdif.beer <- diff(dif.beer, lag=12)

plot(seasdif.beer, main='Seasonal Difference of Log Beer Production in Australia', xlab='Year', ylab='Difference')
```

Plot the ACf and PACF functions of the second differencing.
```{r}
# ACF plot
acf(seasdif.beer, lag.max=36, main = 'ACF for Second Difference of Time-Series')

# PACF plot
pacf(seasdif.beer, lag.max=36, main = 'PACF for Second Difference of Time-Series')
```

Fit a SARIMA model to the data.

non-seasonal: ACF and PACF both decay indicating an ARMA model.

seasonal: ACF and PACF both decay indicating an ARMA model.

There is still seasonality at lag 12 so I will fit a few different SARIMA models and compare AIC values.
```{r}
arima(log.beer, order=c(2,1,3), seasonal=list(order=c(1,1,1), period=12))

arima(log.beer, order=c(2,1,3), seasonal=list(order=c(0,1,1), period=12))

arima(log.beer, order=c(2,1,3), seasonal=list(order=c(0,1,2), period=12))


fit <- arima(log.beer, order=c(2,1,3), seasonal=list(order=c(0,1,1), period=12))
```

The ARIMA(2, 1, 3) x (0, 1, 2) model has the lowest AIC so I'll check the diagnostics on it
```{r}
# plot ACf and PACF
acf(fit$residuals, lag.max=36)
pacf(fit$residuals, lag.max=36)

# check the Ljung Box statistic
tsdiag(fit)

# Check for normality
qqnorm(fit$residuals)
qqline(fit$residuals)
```

The model looks to be a reasonable fit. Forecast using the model:
```{r}
# get predictions from model
pred <- predict(fit, n.ahead=12)

# plot original time-series
plot(ts.beer, main='Monthly Beer Production in Australia', xlab='Year', ylab='Beer (Megaliters)', xlim=c(1956, 1997))

# add forecast predictions coverted back to original scale from log scale
lines(exp(pred$pred), col='red')
```

A closer look at the forecast, with 95% confidence interval
```{r}
# plot original time-series
plot(ts.beer, main='Monthly Beer Production in Australia', xlab='Year', ylab='Beer (Megaliters)', xlim=c(1993, 1996.5))

# add forecast and confidence intervals
lines(exp(pred$pred), col='red')
lines(exp(pred$pred+2*pred$se), col='blue')
lines(exp(pred$pred-2*pred$se), col='blue')
```

Compare to Holt Winters exponential smoothing forecast method
```{r}
hlt.beer <- hw(ts.beer, seasonal='multiplicative', h=12)

plot(hlt.beer)
plot(hlt.beer, xlim=c(1993, 1996.5))
```

















